---
title: "Lab 02 - Plastic waste"
author: "Marcus Minko"
date: "01-20-2022"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

### Exercise 1

You can't tell much from this visual, besides the fact that Oceania and South America have fewer countries registering waste per cap data.

```{r plastic-waste-continent}
ggplot(data = plastic_waste, mapping = aes(x = plastic_waste_per_cap, color = continent)) + geom_histogram() + facet_wrap(~ continent)
```

### Exercise 2

```{r plastic-waste-density}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, 
                     fill = continent)) +
  geom_density(alpha = 0.3)
```

### Exercise 3

The color and fill features are a subordinate part of the mapping = aes 
function, and alpha is subordinate or nested within the geom_density function. These features can only be defined within their appropriate respective functions.

### Exercise 4

With the violin plot, opposed to boxplots, we lose some specificity regarding mean, quartiles, etc.

```{r plastic-waste-violin}
ggplot(data = plastic_waste,    mapping = aes(x = continent, y = plastic_waste_per_cap, color = continent, fill = continent)) + geom_violin()

```

### Exercise 5

Based ont he scatterplot, there is not a clear relationship between the two variables. Perhaps one could say that there is a tendency that they are some what negatively correlated in that as waste per cap goes up, mismanaged waste goes down.

```{r plastic-waste-mismanaged}
ggplot(data = plastic_waste, mapping = aes(x = plastic_waste_per_cap, y = mismanaged_plastic_waste)) + geom_point()
```

### Exercise 6

It seems that Asia is the worst offender of waste mismanagement.

```{r plastic-waste-mismanaged-continent}
ggplot(data = plastic_waste, mapping = aes(x = plastic_waste_per_cap, y = mismanaged_plastic_waste, color = continent, shape = continent)) + geom_point() + scale_color_viridis_d()
```

### Exercise 7

There does not appear to be a linear relationship in either case.

```{r plastic-waste-population-total}
ggplot(data = plastic_waste, mapping = aes(x = plastic_waste_per_cap, y = total_pop, color = continent, shape = continent)) + geom_point() + scale_color_viridis_d()
```

```{r plastic-waste-population-coastal}
ggplot(data = plastic_waste, mapping = aes(x = plastic_waste_per_cap, y = coastal_pop, color = continent, shape = continent)) + geom_point() + scale_color_viridis_d()
```

### Exercise 8

Remove this text, and add your answer for Exercise 8 here.

```{r recreate-viz}
# insert code here
```

## Pro-Tips

### Excercise 3

Try this :D

ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()+
  geom_boxplot(width=.3, fill="green") +
  stat_summary(fun.y=median, geom="point") 
  
### Exercise 5 

Helpful reference:http://www.sthda.com/english/wiki/ggplot2-themes-and-background-colors-the-3-elements



